\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved section "Replication of textbook results" (systemfit_reliability.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@956 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\input{systemfit_reliability.tex}
\input{systemfit_summary.tex}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Replication of textbook results}\label{sec:reliability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{prefix.string=trash/replication}
<<echo=FALSE>>=
options( prompt = "R> " )
@

In this section, we reproduce results from several textbook examples
using the \pkg{systemfit} package for several reasons.
First, a comparison of \pkg{systemfit}'s results
with results published in the literature
confirms the reliability of the \pkg{systemfit} package.
Second, this section helps teachers and students
become familiar with using the \pkg{systemfit} package.
Third, the section encourages reproducible research,
which should be a general goal in scientific analysis
\citep{buckheit95,schwab00}.
For instance, by preparing this section,
the exact estimation methods of the replicated analyses have been discovered
and a few errors in \citet{greene03} have been found
\citep[see][]{greene06a}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Kmenta (1986): Example on p. 685 (food market)]
   {Kmenta (1986): Example on p.~685 (food market)}

First, we reproduce an example taken from \citet[p.~685]{kmenta86}.
The data are available from Table~13-1 (p.~687),
and the results are presented in Table~13-2 (p.~712) of this book.

Before starting the estimation,
we load the data and specify the two formulas of the model
as well as the instrumental variables.
Then the equation system is estimated by OLS, 2SLS, 3SLS, and
iterated 3SLS.
After each estimation, we provide the commands
to print the estimated coefficients.
<<systemfit_replication,echo=FALSE>>=
library( "systemfit" )
@
<<systemfit_replication,results=hide>>=
data( "Kmenta" )
eqDemand <- consump ~ price + income
eqSupply <- consump ~ price + farmPrice + trend
inst <- ~ income + farmPrice + trend
system <- list( demand = eqDemand, supply = eqSupply )
@
OLS estimation:
<<systemfit_replication,results=hide>>=
fitOls <- systemfit( system, data = Kmenta )
round( coef( summary( fitOls ) ), digits = 4 )
@
2SLS estimation:
<<systemfit_replication,results=hide>>=
fit2sls <- systemfit( system, method = "2SLS", inst = inst, data = Kmenta )
round( coef( summary( fit2sls ) ), digits = 4 )
@
3SLS estimation:
<<systemfit_replication,results=hide>>=
fit3sls <- systemfit( system, method = "3SLS", inst = inst, data = Kmenta )
round( coef( summary( fit3sls ) ), digits = 4 )
@
Iterated 3SLS estimation:
<<systemfit_replication,results=hide>>=
fitI3sls <- systemfit( system, method = "3SLS", inst = inst, data = Kmenta,
   maxit = 250 )
round( coef( summary( fitI3sls ) ), digits = 4 )
@

The above commands return exactly the same
coefficients and standard errors as published in \citet[p.~712]{kmenta86}
except for two minor exceptions:
two standard errors of the 2SLS estimation deviate by $0.0001$.
However, this difference is likely due to rounding errors
in \code{systemfit} or \citet{kmenta86} and
is so small that it empirically does not matter.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Greene (2003): Example 15.1 (Klein's model I)}
\label{sec:KleinsModel}

Second, we try to replicate Klein's °Model I° \citep{klein50}
that is described in \citet[p.~381]{greene03}.
The data are available from the online complements to \citet{greene03},
Table~F15.1
(\url{http://pages.stern.nyu.edu/~wgreene/Text/tables/TableF15-1.txt}),
and the estimation results are presented in Table~15.3 (p.~412).

Initially, the data are loaded and three equations
as well as the instrumental variables are specified.
As in the example before, the equation system is estimated by
OLS, 2SLS, 3SLS, and iterated 3SLS, and
commands to print the estimated coefficients are presented.
<<systemfit_replication>>=
data( "KleinI" )
eqConsump  <- consump ~ corpProf + corpProfLag + wages
eqInvest   <- invest ~ corpProf + corpProfLag + capitalLag
eqPrivWage <- privWage ~ gnp + gnpLag + trend
inst <- ~ govExp + taxes + govWage + trend + capitalLag + corpProfLag + gnpLag
system <- list( Consumption = eqConsump, Investment = eqInvest,
   PrivateWages = eqPrivWage )
@
OLS estimation:
<<systemfit_replication,results=hide>>=
kleinOls <- systemfit( system, data = KleinI )
round( coef( summary( kleinOls ) ), digits = 3 )
@
2SLS estimation:
<<systemfit_replication,results=hide>>=
klein2sls <- systemfit( system, method = "2SLS", inst = inst, data = KleinI,
   methodResidCov = "noDfCor" )
round( coef( summary( klein2sls ) ), digits = 3 )
@
3SLS estimation:
<<systemfit_replication,results=hide>>=
klein3sls <- systemfit( system, method = "3SLS", inst = inst, data = KleinI,
   methodResidCov = "noDfCor" )
round( coef( summary( klein3sls ) ), digits = 3 )
@
iterated 3SLS estimation:
<<systemfit_replication,results=hide>>=
kleinI3sls <- systemfit( system, method = "3SLS", inst = inst, data = KleinI,
   methodResidCov = "noDfCor", maxit = 500 )
round( coef( summary( kleinI3sls ) ), digits = 3 )
@

Again, these commands return almost the same results
as published in \citet{greene03}.%
\footnote{
There are two typos in Table~15.3 (p.~412).
Please take a look at the errata
\citep{greene06a}.
}
There are only two minor deviations,
where these values differ merely in the last digit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Greene (2003): Example 14.1 (Grunfeld's investment data)}
\label{sec:grunfeld-greene}

Third, we reproduce Example~14.1 of \citet[p.~340]{greene03}
that is based on \citet{grunfeld58}.
The data are available from the online complements to \citet{greene03},
Table~F13.1
(\url{http://pages.stern.nyu.edu/~wgreene/Text/tables/TableF13-1.txt}).
Several different versions of the °Grunfeld° data set can be found,
whereas the version of \citet{greene03}
is considered incorrect \citep{cummins01}.
However, we use this incorrect version to replicate the results
in \citet{greene03}, Tables~14.1 and~14.2 (p.~351).%
\footnote{
A correct version of this data set with five additional firms
is available as data set \code{Grunfeld} in the \pkg{Ecdat} package
\citep{r-Ecdat-0.1-5}.
}

First, we load the data and the \pkg{plm} package,
indicate the individual (cross-section) and time identifiers,
and specify the formula to be estimated.
Then, the system is estimated by OLS, pooled OLS, SUR, and pooled SUR.
After each estimation, we show the commands to print the estimated coefficients,
the $\sigma^2$ values of the OLS estimations, and
the residual covariance matrix as well as the residual correlation matrix
of the SUR estimations.
<<systemfit_replication>>=
data( "GrunfeldGreene" )
library( "plm" )
GGPanel <- plm.data( GrunfeldGreene, c( "firm", "year" ) )
formulaGrunfeld <- invest ~ value + capital
@
OLS estimation (Table 14.2):
<<systemfit_replication,results=hide>>=
greeneOls <- systemfit( formulaGrunfeld,
   data = GGPanel )
round( coef( summary( greeneOls ) ), digits = 4 )
round( sapply( greeneOls$eq, function(x){return(summary(x)$ssr/20)} ), digits = 3 )
@
pooled OLS (Table 14.2):
<<systemfit_replication,results=hide>>=
greeneOlsPooled <- systemfit( formulaGrunfeld,
   data = GGPanel, pooled = TRUE )
round( coef( summary( greeneOlsPooled$eq[[1]] ) ), digits = 4 ) #$
sum( sapply( greeneOlsPooled$eq, function(x){return(summary(x)$ssr)}) )/100
@
SUR estimation (Table~14.1):
<<systemfit_replication,results=hide>>=
greeneSur <- systemfit( formulaGrunfeld, method = "SUR",
   data = GGPanel, methodResidCov = "noDfCor" )
round( coef( summary( greeneSur ) ), digits = 4 )
round( greeneSur$residCov, digits = 3 ) #$
round( summary( greeneSur )$residCor, digits = 3 ) #$
@
pooled SUR estimation (Table~14.1):
<<systemfit_replication,results=hide>>=
greeneSurPooled <- systemfit( formulaGrunfeld, method = "SUR",
   data = GGPanel, pooled = TRUE, methodResidCov = "noDfCor",
   residCovWeighted = TRUE )
round( coef( summary( greeneSurPooled$eq[[1]] ) ), digits = 4 ) #$

round( greeneSurPooled$residCov, digits = 3 ) #$
round( cov( residuals( greeneSurPooled ) ), digits = 3 )
round( summary( greeneSurPooled )$residCor, digits = 3 ) #$
@

These commands return
nearly the same results as published in \citet{greene03}.%
\footnote{
There are several typos and errors in Table~14.1 (p.~412).
Please take a look at the errata of this book
\citep{greene06a}.
}
We present two different commands to print
the residual covariance matrix of the pooled SUR estimation.
The first calculates the covariance matrix without centering the residuals
(see Section~\ref{sec:residcov});
the returned values are equal to those published in \citet[p.~351]{greene03}.
The second command calculates the residual covariance matrix
after centering the residuals;
these returned values are equal to those published in the errata
\citep{greene06a}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection[Theil (1971): Example on p. 295ff (Grunfeld's investment data)]
   {Theil (1971): Example on p.~295ff (Grunfeld's investment data)}

Finally, we estimate an example taken from \citet[p.~295ff]{theil71}
that is also based on \citet{grunfeld58}.
The data are available in Table~7.1 of \citet[p.~296]{theil71}.
They are a subset of the data set published by \citet{greene03}
(see Section~\ref{sec:grunfeld-greene}).

After extracting the data from the \code{GrunfeldGreene} data set,
the individual (cross-section) and time identifiers are indicated.
Then, the formula is specified,
and the model is estimated by OLS and SUR.
Commands to print the estimated coefficients are reported
after each estimation.
<<systemfit_replication>>=
GrunfeldTheil <- subset( GrunfeldGreene,
   firm %in% c( "General Electric", "Westinghouse" ) )
GTPanel <- plm.data( GrunfeldTheil, c( "firm", "year" ) )
formulaGrunfeld <- invest ~ value + capital
@
OLS estimation (page 295)
<<systemfit_replication,results=hide>>=
theilOls <- systemfit( formulaGrunfeld,
   data = GTPanel )
round( coef( summary( theilOls ) ), digits = 3 )
@
SUR estimation (page 300)
<<systemfit_replication,results=hide>>=
theilSur <- systemfit( formulaGrunfeld, method = "SUR",
   data = GTPanel, methodResidCov = "noDfCor" )
round( coef( summary( theilSur ) ), digits = 3 )
@

These commands
return exactly the same results as published in
\citet[pp.~295, 300]{theil71}.

Now, we apply an $F$ test to check whether the slope parameters are equal
for General Electric and Westinghouse (pages~313--315).
Then we re-estimate the model under these restrictions on the coefficients.

$F$ test (page 313--315)%
\footnote{%
The same restriction can be specified also symbolically by
\code{RMatrix <- c("General.Electric\_value = Westinghouse\_value",
"General.Electric\_capital = Westinghouse\_capital")}
}
<<systemfit_replication,results=hide>>=
RMatrix <- rbind( c( 0, 1, 0, 0, -1, 0 ), c( 0, 0, 1, 0, 0, -1 ) )
linear.hypothesis( theilSur, RMatrix )
@

Restricted SUR estimation (page~316)
<<systemfit_replication,results=hide,keep.source=TRUE>>=
theilSurRestr <- systemfit(formulaGrunfeld, method = "SUR",
   data = GTPanel, methodResidCov = "noDfCor", restrict.matrix = RMatrix,
   residCovRestricted = FALSE)
round(coef(summary(theilSurRestr)), digits = 3)
@

The method \codeD{linear}{hypothesis} returns the same value of the $F$ statistic
as published in \citet[p.~315]{theil71}.
Hence, we arrive at the the same conclusion:
we accept the null hypothesis (restrictions on the coefficients are true)
at the 5~percent significance level.

Also the results of the restricted SUR estimation are identical
to the results published in
\citet[p.~316]{theil71}.

\input{systemfit_summary.tex}
